Practical Machine Learning - Project Summary
============================================

## Overview:
The project aims to predict the manner in which participants did "Weight Lifting Exercise".

## Processing
The first stage is loading the training dataset, spliting it to 2 subsets - one for creating the model and another subset for checking the accuracy of the model.
```{r results='hide' ,message=FALSE, warning=FALSE}
library(caret)
setwd("D://docs//studying//Coursera//ML")
url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
#download.file(url,destfile="pml-training.csv") #no need to download the file when compiling the Knit, since it is already in the workspace
training <-read.csv("pml-training.csv",header=TRUE,na.strings = c("", "NA"))
sampleIndex <- createDataPartition(training$classe,times=1,p=0.75,list=FALSE)
sample <- training[sampleIndex,2:length(names(training))]

featurePlot(x=sample[,1:7],y=sample$classe,plot="pairs")

```

After some attempts to fit a good model and based on the course forum, i came into conclusion to remove the columns with have many missing values and those that are related to the way the experiment was made (all presented in the chart above).

```{r , warning=FALSE}
countNa <- sort(apply(training,2,function(column) {sum(is.na(column))}))
onlyColumnWithoutNa <- countNa[countNa ==0]
columnNames <- names(onlyColumnWithoutNa)
selectedColumns <- columnNames[8:length(columnNames)] # remove X and user name and other attributes that relate to the way the experiment made
sample.selectedColumns <- sample[,selectedColumns]
```
## Modeling

Fit a random forest model with cross validation of 10-fold. I have checked other approach such as boot632 and tried to also run a PCA, yet all the other approach yieled worst results
```{r message=FALSE, , warning=FALSE}
rf_model_cv2<-train(classe~.,data=sample.selectedColumns,method="rf",trControl=trainControl(method="cv"),prox=TRUE)
```
### Model Details:
```{r}
rf_model_cv2
plot(varImp(rf_model_cv2,type=2))
```

The model error rate and the confusion matrix for the 5 "classe"s:
```{r}
rf_model_cv2$finalModel
```

Checking the model with the remaining part of the training data:
```{r}
traingTest <- training[-sampleIndex,]
traingTest.selectedColumns <- traingTest[,selectedColumns]
pred <- predict(rf_model_cv2,traingTest.selectedColumns)
confusionMatrix(traingTest.selectedColumns$classe,pred)
```

## Predicting the classe attribute of the testing data
```{r}
testFileName <-"pml-testing.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
#download.file(testUrl,destfile=testFileName)
testing <-read.csv(testFileName,header=TRUE,na.strings = c("", "NA"))
selectedColumnsWithoutClasse <- selectedColumns[1:length(selectedColumns)-1]
testingData <- testing[,selectedColumnsWithoutClasse]
predTesting <- predict(rf_model_cv2,testingData)
```

The 20 predicted values:
```{r}
predTesting
```

## Summary:
The random forest algorithm yields a high predication accuracy and a low error rate. The RF includes a cross validation check.
